Methods to prioritise pop-up active transport infrastructure

### 2020-06-18 article.Rmd
Reduce paper compile time

```{r load}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("plot")
rtid = readr::read_csv("rtid.csv")
region = ukboundaries::leeds
```

```{r preprocess}
# remove motorways
r_original$highway_type = r_original$highway
r_original$highway_type = gsub(pattern = "_link", replacement = "", r_original$highway)
highway_table = table(r_original$highway)
highway_rare = highway_table[highway_table < nrow(r_original) / 100]
highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|ped", x = names(highway_rare))]
r_cleaned = r_original %>% 
  filter(!grepl(pattern = "motorway", x = highway)) %>% 
  mutate(highway_type = case_when(
  highway_type %in% highway_remove ~ "other",
  grepl(pattern = "path|track|trunk", highway_type) ~ "other",
  grepl(pattern = "ped|liv", highway_type) ~ "pedestrian/living_street",
  TRUE ~ highway_type
  )) %>% 
  mutate(maxspeed = case_when(
    maxspeed <= 20 ~ "20 mph or less",
    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
    maxspeed > 30 ~ "40+ mph",
  )) %>% 
  mutate(cycling_potential = as.numeric(pctgov)) %>% 
  select(name, ref, highway, maxspeed, pctgov, width, Existing, length, highway_type, cycling_potential, rtid, idGlobal)
# table(r_cleaned$maxspeed)
rj = inner_join(r_cleaned, rtid) %>% 
  mutate(lanes_f = abs(lanespsvforward) + abs(lanesforward)) %>% 
  mutate(lanes_b = abs(lanespsvbackward) + abs(lanesbackward)) %>% 
  mutate(n_lanes_numeric = lanes_f + lanes_b) %>% 
  mutate(n_lanes = as.character(n_lanes_numeric))
rj$n_lanes_numeric[rj$n_lanes_numeric <= 0] = 1 
rj$n_lanes[rj$n_lanes == "0"] = "1"
rj$n_lanes[rj$n_lanes == "4"] = "4+"
rj$n_lanes[rj$n_lanes == "5"] = "4+"
rj$n_lanes[rj$n_lanes == "6"] = "4+"
cy = r_cleaned %>% filter(highway == "cycleway")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
```

@@

```{r gsub, fig.cap="Illustration of geographic subsetting based on distance to a central point (Leeds city centre in this case) rather than based on location within somewhat arbitrarily shaped city boundaries. Radii of 5 km, 8 km and 10 km are shown for reference (note that some roads within 10 km of the center are outside the regional boundary)."}
# nrow(r_central)
# nrow(rj)
square = stplanr::geo_bb(shp = city_centre_buffer)
m1 = tm_shape(r_main_region) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(region) + tm_borders() +
  tm_shape(square) + tm_borders(lty = 3) +
  tm_scale_bar() 
# m1
m2 = tm_shape(r_main) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(city_centre_buffer_small) + tm_borders(lty = 2) +
  tm_shape(city_centre_buffer_large) + tm_borders(lty = 2) +
  tm_shape(city_centre_buffer) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.show = FALSE)
tmap_arrange(m1, m2, nrow = 1)
```

@@

touching_list = st_touches(r_pct_lanes_all)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes_all$group = components$membership
r_pct_lanes = r_pct_lanes_all %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = weighted.mean(cycling_potential, w = length, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)

>>>>

```{r load}
library(sf)
library(tidyverse)
library(tmap)
region_name = "West Yorkshire"
# source("code/build.R")
# saveRDS(r_main_region, "r_main_region.Rds")
# r_main_leeds = r_main_region[region, ]
# saveRDS(r_main_leeds, "r_main_leeds.Rds")
# piggyback::pb_upload("r_main_leeds.Rds")
if(!file.exists("r_main_leeds.Rds")) {
  u = "https://github.com/cyipt/popupCycleways/releases/download/v1/r_main_leeds.Rds"
  download.file(u, "r_main_leeds.Rds")
}
rj = readRDS("r_main_leeds.Rds")
region = ukboundaries::leeds
tmap_mode("plot")
```

```{r parameters}
regexclude = "welling"
min_cycling_potential = 0
min_grouped_cycling_potential = 100
min_grouped_length = 500
city_centre_buffer_radius_small = 5000
city_centre_buffer_radius = 8000
city_centre_buffer_radius_large = 10000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer_small = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_small)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
city_centre_buffer_large = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_large)
```

```{r preprocess}
cy = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/cycleways.geojson")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
```

@@

```{r gsub, fig.cap="Illustration of geographic subsetting based on distance to a central point (Leeds city centre in this case) rather than based on location within somewhat arbitrarily shaped city boundaries. Radii of 5 km, 8 km and 10 km are shown for reference (note that some roads within 10 km of the center are outside the regional boundary)."}
# nrow(r_central)
# nrow(rj)
square = stplanr::geo_bb(shp = city_centre_buffer)
m1 = tm_shape(r_main_region) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(region) + tm_borders(col = "black", lwd = 2) +
  tm_shape(square) + tm_borders(lty = 3) +
  tm_scale_bar() 
# m1
m2 = tm_shape(r_main) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(city_centre_buffer_small) + tm_borders(lty = 2) +
  tm_shape(city_centre_buffer_large) + tm_borders(lty = 2) +
  tm_shape(region) + tm_borders(col = "black", lwd = 2) +
  tm_shape(city_centre_buffer) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.show = FALSE)
tmap_arrange(m1, m2, nrow = 1)
```

@@

touching_list = st_touches(r_pct_lanes_all)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes_all$group = components$membership
r_pct_lanes = r_pct_lanes_all %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = weighted.mean(cycling_potential, w = length, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length) %>% 
  ungroup()
